---
title: "10xVisualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo = FALSE, message = FALSE, eval = TRUE}
library(BiocManager)
library(SingleCellExperiment)
library(purrr)
library(DropletUtils)
library(keras)
library(scran)
library(scater)
library(r10xDL)
library(umap)
```

First we will read and subset the data into manageable sizes. 

```{r loadData, echo = TRUE}
subsetSize <- 500

sce1 <- read10xCounts("../10Xdata/vdj_v1_hs_aggregated_donor1_filtered_feature_bc_matrix.h5")
numRows <- nrow(counts(sce1))
sce1 <- sce1[1:numRows, 1:subsetSize]

sce2 <- read10xCounts("../10Xdata/vdj_v1_hs_aggregated_donor2_filtered_feature_bc_matrix.h5")
sce2 <- sce2[1:numRows, 1:subsetSize]

sce3 <- read10xCounts("../10Xdata/vdj_v1_hs_aggregated_donor3_filtered_feature_bc_matrix.h5")
sce3 <- sce3[1:numRows, 1:subsetSize]

sce4 <- read10xCounts("../10Xdata/vdj_v1_hs_aggregated_donor4_filtered_feature_bc_matrix.h5")
sce4 <- sce4[1:numRows, 1:subsetSize]
```

Separate the matrix into separate segments by type of observation, where gex1 is all gene expression rows, prtn1 is all protein rows, and ag1 as all antigen rows.
```{r separateRows, echo = TRUE}
gex1 <- sce1[rowData(sce1)$Type == "Gene Expression"]
ngex1 <- sce1[rowData(sce1)$Type == "Antibody Capture"] #gets all data that is not gene expression data -> antibody capture data
prtn1 <- ngex1[1:14,] #dirty way to get all protein data
ag1 <- ngex1[15:64,] #dirty way to get all antigen data
```

## Antigen Data

To show a histogram of each antigen's frequency: (code not run)
```{r antibodyFrequencyHistograms, echo = TRUE, eval = FALSE}
for (i in 1:50)
{
  ag <- ag1[i,]
  agList <- c(unname(as.matrix(counts(ag))))
  agList <- agList[agList > 0] #excludes cells with no measured antigen
  hist(agList, breaks = 50, main = rownames(ag1)[i])
}
```

Create a vector of what  each cell produces, graph the results, and append it to coldata. (code not run)
```{r determineAgType, echo = TRUE, eval = FALSE}
countsag1 <- counts(ag1)
getAntigen <- function(i) {
  cellAntigens = c(unname(as.matrix(countsag1[,i])))
  indexMax = which.max(cellAntigens)
  if (indexMax != 1)
    rownames(ag1)[indexMax]
  else
    cellAntigens[1] = cellAntigens[1] - 1
    if (which.max(cellAntigens) == 1)
      rownames(ag1)[1]
    else
      rownames(ag1)[which.max(cellAntigens)]
}
agRow <- unlist(lapply(1:subsetSize, getAntigen)) #agRow is a list of characters, showing the most prevalent antigen in each cell
agVector <- c(unname(as.matrix(agRow)))
agTable <- table(agVector)
par(mar = c(10, 4, 4, 2) + 0.1)
par(cex.axis = .5)
barplot(agTable, main = "Distribution of Antigen Frequencies", las = 2)
par(mar = c(5, 4, 4, 2) + 0.1)
par(cex.axis = 1)

#colData(sce1)$agType <- agVector
colData(gex1)$agType <- agVector
```

We will prep the data by normalizing and running a PCA. Currently I am feeding it GEX data because the antigen data is too small and too sparse to function correctly. 

```{r graphPrep, echo = TRUE, eval=TRUE}
sce <- sce1 #change this
#sce <- cleanSCE(sce)
counts(sce) <- as.matrix(counts(sce)) #turn the matrix from a delayed matrix to a regular matrix so that scran can operate on it
sce <- scran::computeSumFactors(sce)
sce <- scater::normalize(sce)
sce <- scater::runPCA(sce)
```

Here is to graph the PCA of GEX data.

```{r proteinGraph, echo = TRUE, eval=TRUE}
plotReducedDim(sce, use_dimred = "PCA")
```

If you wanted to graph the GEX data for each antigen: (code not run)

```{r redDimByAGType, echo = TRUE, eval= FALSE}
getPlot <- function(x) {
  sceSubset = sce[,(sce$agType == rowData(ag1)$ID[x])]
  plotReducedDim(sceSubset, use_dimred = "PCA", colour_by = "agType")
}
plots <- lapply(1:50, getPlot)
plots
```

## Protein data

To print out a UMAP of the protein data.

```{r UMAP, echo = TRUE}

prtnMatrix <- t(data.frame(log(1+counts(ngex1)))) #extract the delayedmatrix of data and convert it to the right data frame
prtnLayout <- umap(prtnMatrix)$layout #gets the layout portion of the UMAP
ggplot(data.frame(prtnLayout)) + geom_point(aes(prtnLayout[,1], prtnLayout[,2])) #maps the first 2 dimensions of the prtn UMAP
```

For each protein type, we will plot a graph where our data points are colored in accordance with that protein's level of expression

```{r UMAPbyProteinCount, echo = TRUE}
prtnLayout <- prtnLayout
getGGPlot <- function(i) { #i is the numbered protein we want to see
  if (any(counts(prtn1[i,]) > 0)) {
    prtnAugmented <- cbind(prtnLayout, counts(prtn1)[i,])
    ggplot(data.frame(prtnAugmented),  aes(prtnAugmented[,1], prtnAugmented[,2], colour = prtnAugmented[,3])) +
      geom_point() +
      scale_colour_gradientn(colours = rainbow(4)) +
      ggtitle(rowData(prtn1)$ID[i])
  }
  else {
    NULL
  }
}
ggPlots <- lapply(1:nrow(prtn1), getGGPlot)
ggPlots <- Filter(Negate(is.null), ggPlots)
ggPlots
```






